\begin{abstract}
We introduce \ours, a data-free method for extracting a valence direction from the
residual stream of any instruction-tuned large language model using just nine
LLM-generated stories per emotion class (81 samples total). The method is
\emph{trivially simple}: generate nine short paragraphs for each of nine emotion
categories, extract their final-layer hidden states, average per class to form
nine centroids, and take the first principal component of the mean-centered
centroids as the valence axis. We show that this single 1\,536-dimensional
direction, computed without any labeled training data, is competitive with
supervised classifiers across a wide range of tasks: it achieves AUC \sstauc{}
(\sstacc{} accuracy with median thresholding) on SST-2, AUC \imdbauc{} on IMDB,
AUC \yelpauc{} on Yelp Polarity, AUC \twauc{} on Twitter SemEval-17, correlates
\huliur{} with the Hu \& Liu 6{,}789-word sentiment lexicon, and \warrinerr{}
with the continuous Warriner \textit{et al.\@} 2013 valence lexicon. A
data-efficiency curve shows that a single story per class
(9 total samples) already gives AUC 0.74 on SST-2, and that nine stories per
class match the accuracy of a supervised logistic regression trained on
$\sim\!80$--$100$ labeled SST-2 examples --- a $\sim\!10$--$30\times$
data-efficiency gain over supervised training. Layer-wise probing reveals a striking
``V-shape'' pattern: the valence direction appears strongly at layer~1 (inherited
from the embedding layer), fades to near zero through the context-integration
zone (layers 2--26), then abruptly \emph{re-crystallizes} at the final two
blocks. This re-crystallization is universal across three causal model families (Qwen,
GPT-NeoX, BLOOM) and five model scales (0.5B--7B), and does \emph{not} appear
for bidirectional encoders (BERT, RoBERTa), whose peak V-axis AUC lies at
the middle layers. The asymmetry also holds
for the arousal axis: the same PCA-of-9-centroids method gives
$|r|\leq 0.16$ for arousal at every layer, every model, and every test we ran. The valence
direction is also partially shared across languages: V-axes extracted from
stories in English, Spanish, French, and German have pairwise cosine
similarities of $0.16$--$0.47$, while the arousal axes are essentially
orthogonal. Probing is \speedup{} faster than direct prompting at inference,
produces continuous scores useful for ranking, and reveals a novel
interpretability insight about how LLMs represent affect. Our findings
characterize \emph{when} and \emph{where} late-layer linear probing recovers
usable concept directions, and provide a method that requires zero labels,
zero gradient updates, and zero prompt engineering.
\end{abstract}
