\section{Discussion}
\label{sec:discussion}

\subsection{What does the V-shape tell us about LLM computation?}

The striking observation that the V-axis at layer 14 is nearly orthogonal
(cosine $0.067$) to the V-axis at layer 28, while layer 27 is 95\% aligned
with layer 28, suggests a specific computational story.

Early layers (L1--L3) inherit a \emph{word-level} valence signal from the
token embeddings. Word embeddings like ``happy'' and ``sad'' are already
far apart in embedding space, a well-known property of distributed lexical
representations \cite{mikolov2013linguistic, bolukbasi2016man}. Probing at
L1 recovers this signal at AUC $0.75$ on SST-2.

Middle layers (L4--L26) perform context integration. During this phase, the
word-level valence signal is mixed with syntactic, discourse, and task-relevant
features, and the \emph{linear} valence direction is not preserved. A PCA
probe at any middle layer recovers essentially nothing (AUC $\approx 0.53$).

Final layers (L27--L28) prepare the residual stream for the LM head's
next-token prediction. For causal language modeling, the LM head must
produce coherent next-word distributions, which requires the residual stream
to expose ``decodable'' semantic directions. We interpret the final-layer
re-crystallisation as a re-linearisation of task-relevant features: valence
re-emerges as an approximately linear direction that the LM head can read
off. This interpretation is consistent with the linear representation
hypothesis \cite{park2024linear}, which predicts that linear directions
should be observable \emph{in the subspace the LM head acts on} --- i.e.,
the final residual stream --- not necessarily throughout the network.

This view makes a testable prediction: non-causal models
(masked or bidirectional models, which do not share the LM-head constraint
in the same way) should show a different layer-wise pattern. Our
supplementary BERT/RoBERTa measurements are consistent with this
prediction: encoder V-axis AUC peaks at middle layers, not at the final
layer.

\subsection{Cross-language robustness and real-world applicability}

The cross-lingual transfer result (Table~\ref{tab:crosslang_transfer}) is
the strongest evidence that \ours{} is useful in practice. In low-resource
or rapid-prototyping deployments, sentiment labels may not exist in the
target language at all. Our measurements show that a zero-label English
axis, extracted once from 81 LLM-generated stories, classifies held-out
stories in Spanish, French, German, and Chinese at AUC $\geq 0.949$ without
translation, re-extraction, or any target-language labels. For settings
that require a single probe deployable across many languages --- global
product feedback monitoring, multilingual content safety triage, rapid
A/B testing of marketing copy --- this is a substantially lower-cost entry
point than per-language supervised fine-tuning. The method's
data-efficiency curve (one story per class already beats supervised
$N=10$) additionally makes it attractive for rapid iteration: a
domain-specific V-axis can be regenerated in seconds on a V100.

\subsection{When does \ours{} work?}

Consolidating results across Sections~\ref{sec:experiments} and
\ref{sec:scope}, \ours{} is most useful when:

\begin{itemize}[leftmargin=*]
\item The target concept has a clean global-polarity structure that the
LLM's final layer re-linearises (valence yes; formality yes on
self-LOOCV; toxicity only partially, because toxicity depends on local
lexical triggers rather than a global scene-level polarity).
\item The practitioner wants a \emph{compressed, interpretable} probe:
a single 1\,536-dim direction whose projection is continuous, ranking-friendly,
and easy to audit.
\item Labels are scarce or unavailable in the target language or domain.
The cross-lingual transfer result and the $N\!=\!100$-label equivalence
quantify the savings.
\item Inference speed matters (\speedup{} faster than direct prompting at
SST-2, with the advantage growing in batched settings).
\end{itemize}

Conversely, the method should not be expected to dominate in two
regimes --- both of which we quantify rather than hide:
(i)~when raw classification accuracy is the primary metric, a full-rank
supervised probe or direct few-shot prompting outperforms us by
$0.04$--$0.15$ (Tables~\ref{tab:sst2} and \ref{tab:counterfactual});
(ii)~when the concept is not a global polarity (arousal, toxicity), the
9-centroid PCA does not reliably recover a usable direction.

\subsection{Valence vs.\ arousal: a scope statement}

The valence/arousal asymmetry defines the regime of the method. Across 5
model families, 5 languages, and every layer tested, the same PCA-of-9-centroids
recipe gives $|r|=0.50$ for valence and $|r|\leq 0.16$ for arousal. Three
mechanisms are compatible with this pattern:

\begin{enumerate}[leftmargin=*]
\item \textbf{Distributed encoding.} Arousal may be encoded in a
non-linear or multi-dimensional manner in LLM residual streams; an
MLP probe or a multi-direction extraction could succeed where a single
linear direction does not. The PC2/PC3 analysis in
Section~\ref{sec:principal_angle} suggests neither PC2 nor PC3 carries
a usable arousal signal, so if it is linear it lives on a later PC
that our small centroid matrix cannot expose.
\item \textbf{Embedding-only arousal.} Arousal words (``calm'',
``frantic'') are separable in embedding space, but the information is
not re-crystallised in late layers the way valence is. This is
consistent with \cite{park2024linear}: LM-head-relevant directions
re-appear at the top; other directions may not.
\item \textbf{Scene-level vs.\ signal-level arousal.} Our 9 emotion
categories span different \emph{valences} but limited \emph{arousal
range}. A re-designed category set explicitly sweeping arousal
(e.g.\ contentment vs.\ excitement, both positive but different
arousal) might expose the arousal axis.
\end{enumerate}

Hypothesis~2 is most consistent with the layer-wise observation that
valence re-crystallises at L27--L28 while arousal does not; hypothesis~3
suggests a clean future-work extension. Rather than treat arousal as a
failure, we treat it as a falsifiable scope statement: the method's
regime of applicability is \emph{valence-like global concepts whose
linear directions survive the late-layer re-linearisation}.

\subsection{Relation to the Linear Representation Hypothesis}

Park, Choe and Veitch \cite{park2024linear} derive geometric tests for
when concepts live on low-dimensional linear subspaces of LLM hidden
states. \ours{} can be viewed as an extraction recipe matching that
predicate: the 9 category centroids form a one-dimensional affine
slice (PC1 captures $55.8\%$ of centroid variance, PC2 $15.2\%$),
and the principal angle to a random direction is $88.8^\circ$,
well-separated from the PC1 direction itself. Our empirical result ---
that PC1 of nine centroids is a cross-task, cross-lingual sentiment
classifier --- is a direct practical instantiation of the linear
representation hypothesis for sentence-level valence.

\subsection{Theoretical Connection to Few-Shot Learning}

\ours{} is a degenerate case of few-shot prototype learning
\cite{snell2017prototypical} where (i)~the prototypes are LLM-generated
rather than human-labeled, (ii)~the query representation is the LLM's own
final layer, (iii)~the classifier is a single projection (not a soft-max
over distances). The key insight is that the LLM itself can serve as both
the data source and the representation source; no human labels are needed
for either role.
