\section{Conclusion}
\label{sec:conclusion}

We presented \ours{}, a fully unsupervised recipe that extracts a linear
valence direction from LLM residual streams using nine LLM-generated
stories per emotion class. The method uses zero human labels, zero
gradient updates, and no operator-specified contrastive pairs. A single
81-story axis transfers across seventeen zero-label evaluations: five
English sentiment benchmarks (SST-2 AUC $0.868$, IMDB $0.895$, Yelp
$0.953$, Rotten Tomatoes $0.843$, Twitter $0.923$), four non-English
sentiment-transfer tests (ES $0.949$, FR $0.991$, DE $0.972$, ZH $1.000$),
three word-valence lexicons ($|r|=0.76, 0.69, 0.95$), and one continuous
VAD regression ($|r|=0.50$). A supervised logistic regression on the same
features needs $N\!\approx\!100$ SST-2 labels to match our zero-label
accuracy.

Three properties define the method's practical value. First, it is robust
across languages: the English axis transfers to four other languages at
AUC $\geq 0.949$ without translation or re-extraction, making it a
deployment-ready probe for multilingual pipelines. Second, it is simpler
than the standard representation-engineering recipe: PCA of nine
category-centroids outperforms the mean-difference contrastive axis of
\cite{zou2023representation, rimsky2024steering, arditi2024refusal} by
$+0.115$ AUC on matched Qwen-1.5B features, while removing the need to
specify binary positive/negative pairs. Third, it is interpretable: the
counterfactual ablation quantifies precisely what the axis captures
(full-rank LogReg reaches AUC $0.911$; removing our axis changes that by
$0.001$), confirming that the axis is the most useful one-dimensional
summary of a richly encoded valence signal, not a causal bottleneck.

On the interpretability side, we identified a ``V-shape'' layer-wise
emergence pattern: the valence direction is present at layer 1
(embedding-level), fades through the middle 24 layers, then
re-crystallises at the final 1--2 transformer blocks. The final-layer
V-axis is nearly orthogonal to the middle-layer V-axis (cosine $0.067$),
showing that the valence direction is freshly constructed just before the
LM head. This pattern replicates across Qwen, Pythia, and Bloom model
families and is consistent with the linear representation hypothesis
\cite{park2024linear}: linear directions should re-appear in the subspace
the LM head acts on.

We define the scope of \ours{} explicitly. The method works on
valence-like global concepts aligned with the linear representation
hypothesis; it does not recover arousal (PCA of the second component
gives $|r|\leq 0.16$ at every layer) and does not fully recover toxicity
(AUC $0.59$ on Jigsaw, where lexical triggers dominate over global
polarity). These bounds are scope statements, not failures: they
delineate the class of concepts for which one-dimensional unsupervised
extraction is expected to succeed.

We hope this work motivates further study of \emph{when and where} LLMs
encode concepts as linear directions, and provides a simple, reproducible,
deployment-oriented baseline for zero-label concept probing in future
interpretability research.
