\begin{abstract}
\label{sec:abstract-sec}
Recent interpretability research reveals that large language models develop internal emotion representations organized along the same valence--arousal dimensions found in human psychology.
We introduce \textbf{\ours}, which exploits this brain--LLM alignment: we extract emotion vectors from instruction-tuned LLMs and use their relational geometry as soft-label knowledge distillation targets for EEG emotion classification, teaching the model that \emph{anger is closer to disgust than to joy} rather than treating all misclassifications equally.
On FACED (9-class, 123 subjects), \ours achieves \(\mathbf{0.628}\) balanced accuracy, improving over the previous state of the art (REVE, NeurIPS 2025) by \(\mathbf{+11.2\%}\) relative, with cross-subject generalization confirmed at $p < 0.001$.
We validate the framework along three orthogonal axes: (i) \emph{cross-backbone}, where KD significantly improves both CBraMod (\(\mathbf{+5.2\%}\), $p < 10^{-4}$) and LaBraM (\(\mathbf{+2.6\%}\), $p = 0.025$); (ii) \emph{cross-dataset}, where with sufficient temporal granularity KD also significantly improves SEED-V on LaBraM (\(\mathbf{+1.85\%}\), $p = 0.037$); and (iii) \emph{cross-LLM-family}, where 11 instruction-tuned models from Qwen2.5 and Gemma-4 produce nearly identical emotion geometries (within-family RSA $\rho \approx 0.89$, cross-family $\rho = 0.82$), with linear cluster purity reaching $100\%$ for positive vs.\ negative valence.
Through Centered Kernel Alignment (CKA), we show that KD increases the similarity between EEG features and \emph{every} LLM in our suite, including LLMs from a different family from the training teacher, indicating that KD absorbs a universal emotion structure rather than overfitting to one specific LLM.
A key architectural finding is that freezing the pretrained backbone and adding lightweight adapters (310K parameters) matches full fine-tuning (4.9M parameters): the critical factor is the \emph{quality of supervision}, not model capacity.
Beyond classification, we show that KD training shapes the EEG feature manifold: cluster separation improves by 38\% (Davies-Bouldin index), and the number of valence-correlated principal components doubles (4 vs.\ 2), enabling data-driven emotion steering that shifts 92\% of samples via contrastive activation addition. In the practical cross-subject setting, this structured manifold enables faster calibration for new BCI users: with just 1 labeled sample per class, KD achieves $+5.7\%$ over baseline, reducing clinical calibration requirements.
We release our code, emotion vectors, and pretrained adapters to advance brain--language alignment research for EEG.
\end{abstract}
