\section{Related Work}
\label{sec:related}
\vspace{-1mm}

\paragraph{EEG Foundation Models.}
The development of large-scale pretrained models for EEG has accelerated rapidly. LaBraM~\cite{jiang2024labram} introduced the first large-scale EEG foundation model, pretraining on $\sim$2{,}500 hours of data using vector-quantized neural spectrum prediction. CBraMod~\cite{wang2025cbramod} proposed a criss-cross transformer that factorizes spatial and temporal attention, pretrained on the Temple University Hospital EEG Corpus (TUEG). REVE~\cite{elouahidi2025reve} scaled pretraining to 60{,}000 hours across 25{,}000 subjects with a novel 4D positional encoding. CoMET~\cite{li2025comet} combined contrastive learning with masked autoencoding. EEG-DINO~\cite{wang2025eegdino} applied hierarchical self-distillation. mdJPT~\cite{mdjpt2025} explored multi-dataset joint pretraining. EEG-GraphAdapter~\cite{eeg_graphadapter2024} proposed graph-based adapter PEFT. These models achieve state-of-the-art performance across diverse EEG tasks but treat emotion recognition as a standard classification problem with integer labels. In contrast, \ours\ leverages the relational structure among emotions as an additional supervisory signal, complementing rather than replacing the pretrained representations. We additionally provide the most thorough cross-backbone validation to date by replicating our framework on both CBraMod and LaBraM and characterizing when KD does and does not generalize across architectures.

\paragraph{Emotion Representations in LLMs.}
Recent interpretability research has revealed that LLMs develop structured internal representations of emotion concepts. Sofroniew et al.~\cite{sofroniew2026anthropic} demonstrated that Claude's emotion vectors organize along valence--arousal axes, are causally functional (driving behavior such as reward hacking), and persist across diverse contexts. Tak et al.~\cite{tak2025emotion_interpretability} showed through probing and activation patching that emotion processing concentrates in mid-layer multi-head self-attention, with peak encoding at 50--75\% depth. Layer-wise probing of LLMs for emotion has revealed depth-dependent encoding of affective information~\cite{decoding_emotion_deep2024}, and controllable generation work has demonstrated that emotion vectors can steer text output~\cite{emotion_vectors_controllable2025}. Brain--LLM alignment studies find that scaling, but not instruction tuning beyond a threshold, increases representational correspondence with neural data~\cite{scaling_brain_alignment2024}. We extend this line of work in two ways. First, we show that the emotion geometry is genuinely universal across instruction-tuned LLM families: 11 models from Qwen2.5 (0.5B--72B) and Gemma-4 (E2B--31B) produce nearly identical RDMs (within-family $\rho \approx 0.89$, cross-family $\rho = 0.82$). Second, we exploit this geometry as a teaching signal for EEG models and show via CKA that the resulting EEG features become \emph{simultaneously} aligned with all 11 LLMs---not just the specific teacher used during training.

\paragraph{EEG Emotion Recognition with Language Models.}
Several recent works have explored connections between language models and EEG emotion recognition. EmotionCLIP~\cite{emotionclip2025} reformulates emotion recognition as EEG-text matching using frozen CLIP text encoders with prompt templates. EMOD~\cite{emod2026} projects discrete emotion labels into a continuous valence--arousal (V--A) space and applies soft-weighted supervised contrastive learning, pretraining on 8 EEG datasets. E$^2$-LLM~\cite{ma2026e2llm} directly integrates an EEG encoder with Qwen-based LLMs through multi-stage training including chain-of-thought reasoning. Our approach differs fundamentally: rather than using LLM text embeddings of emotion \emph{labels} (EmotionCLIP) or predefined V--A coordinates (EMOD), we extract emotion vectors from LLM \emph{hidden states} during emotion-laden story generation, capturing richer relational structure that emerges from the model's internal representations.

\paragraph{Knowledge Distillation and Cross-Modal Transfer.}
Knowledge distillation~\cite{hinton2015distilling} has been applied to EEG primarily for model compression---transferring knowledge from high-density to low-density electrode configurations. Cross-modal KD, where teacher and student operate on different modalities, is less explored for EEG. CMCRD~\cite{cmcrd2025} proposed cross-modal contrastive representation distillation. ViTKD~\cite{vitkd2022} provides practical guidelines for distilling into vision transformers, and our framework benefits from similar lessons (e.g., AMP/fp16 instability when distilling with sharp teacher distributions). In our setting, the LLM teacher provides a \emph{relational} supervision signal (similarity structure among emotion classes) rather than transferring a specific representation, making our approach closer to relational knowledge distillation than standard feature-matching KD.

\paragraph{Representational Alignment via CKA and RSA.}
Centered Kernel Alignment~\cite{kornblith2019cka} and representational similarity analysis are now standard tools for measuring cross-model representation alignment, including in brain--LLM alignment work~\cite{brain_llm_compression2023,scaling_brain_alignment2024} and recent KD analyses~\cite{rethinking_cka_kd2024}. We adopt these as the primary metrics for evaluating whether \ours\ genuinely makes EEG features more LLM-aligned, finding that KD increases linear CKA between EEG features and \emph{every} LLM in our 11-model suite, with the strongest post-KD alignment achieved by LLMs from a different family from the training teacher.

\paragraph{EEG Channel Selection and Montage.}
A long-standing question in EEG analysis is how the choice of electrode montage interacts with downstream learning. The standard 10--20 system~\cite{ten_twenty_montage} provides 19 electrodes with broad coverage, while denser montages such as 10--10 (62 electrodes) provide finer spatial resolution at the cost of higher correlation between adjacent channels. Our cross-dataset analysis shows that the 19-channel 10--20 subset of SEED-V combined with longer 5\,s temporal windows provides a sufficient number of spatial-temporal tokens for KD to take effect.

\paragraph{Parameter-Efficient Fine-Tuning for Neural Signals.}
PEFT methods including adapters~\cite{houlsby2019adapters} and LoRA~\cite{hu2022lora} have transformed NLP and vision fine-tuning but remain underexplored for EEG foundation models. We systematically compare bottleneck adapters, LoRA, and FiLM conditioning~\cite{perez2018film}, finding that adapters with frozen backbone provide both parameter efficiency and a regularization benefit on FACED, matching full fine-tuning with $20\times$ fewer trainable backbone parameters.
