\section{Conclusion}
\label{sec:conclusion}
\vspace{-1mm}

We introduced \ours, a framework that bridges LLM emotion representations and EEG through soft-label knowledge distillation. By extracting emotion vectors from instruction-tuned LLMs and using their similarity structure as soft supervision targets, we achieve state-of-the-art results on FACED ($\mathbf{0.628}$ balanced accuracy, $+11.2\%$ over REVE) and significant improvements on SEED-V, with cross-subject generalization confirmed at $p < 0.001$.

We validated the framework along three orthogonal axes. (i) \textbf{Cross-backbone:} KD significantly improves both CBraMod ($+5.5\%$, $p < 10^{-4}$) and LaBraM ($+2.6\%$, $p = 0.025$), confirming architecture-agnostic benefit. (ii) \textbf{Cross-dataset:} after diagnosing why KD initially appears to fail on SEED-V---the \emph{token-count requirement}---we restored the KD effect on SEED-V via 5\,s preprocessing, achieving $+1.85\%$ on LaBraM ($p = 0.037$) and a positive bootstrap CI on CBraMod. (iii) \textbf{Cross-LLM-family:} 11 instruction-tuned models from Qwen2.5 and Gemma-4 produce nearly identical emotion geometries (within-family $\rho \approx 0.89$, cross-family $\rho = 0.82$), with $100\%$ binary cluster purity for positive vs.\ negative valence across all sufficiently capable instruction-tuned LLMs.

A central representational finding is that KD increases linear CKA between EEG features and \emph{every} LLM in our 11-model suite, including LLMs from a different family from the training teacher, with the highest post-KD alignment achieved by Gemma-4 31B (despite training on Qwen-14B). This argues that KD absorbs a universal emotion structure rather than overfitting to one specific teacher. We additionally find: (1) freezing the pretrained backbone and training lightweight adapters matches full fine-tuning on FACED, suggesting that pretrained EEG features are already rich enough; (2) temporal jitter is the single critical augmentation, with all others dispensable; (3) emotion-specific SFT of the teacher LLM degrades vector quality, suggesting that the useful geometry emerges from broad language modeling.

\noindent\textbf{Limitations and honest negative findings.} We report two negative findings that should temper enthusiasm for naive recipe transfer. \emph{First}, the adapter-PEFT recipe that yields the FACED SOTA degrades performance on SEED-V on \emph{both} backbones tested, indicating that PEFT is dataset-specific and depends on class structure and channel topology. \emph{Second}, the dense 62-channel SEED-V montage is anomalously KD-resistant relative to the 19-channel 10--20 subset---suggesting that channel selection interacts with KD in non-trivial ways. We were also unable to load Llama-4 Scout 17B$\times$16E into our pipeline due to a compatibility issue between bitsandbytes 4-bit quantization, accelerate's hooks, and Llama-4's MoE architecture; this is a software-engineering limitation rather than a scientific finding, but we report it for transparency.

\noindent\textbf{Future Work.} Extending \ours\ to additional EEG foundation models and emotion datasets would further validate its generality. Investigating online adaptation of emotion vectors during training, rather than using fixed pre-extracted vectors, is a promising direction for personalized emotion recognition. Understanding why the dense 62-channel montage resists KD is an open question with implications for EEG preprocessing pipelines beyond emotion classification.

\noindent{\textbf{Data Availability Statement.}}
Our code, pretrained adapter weights, LLM emotion vectors for both FACED and SEED-V (across all 11 LLM models we tested), and the SEED-V 5\,s preprocessing pipeline will be made publicly available upon publication. The code will be released on GitHub.
