\begin{abstract}
Deep ensembling boosts accuracy and calibration by averaging models trained
with different random seeds. We show that a second, essentially free
diversity axis has been hiding in plain sight: the \emph{length} of the
cosine schedule. Training the same architecture for two different epoch
budgets (e.g.\ 100 and 150) with the same seed, optimizer, and
schedule---changing only one integer---reaches \emph{identical} mean
accuracy (within $<\!0.001$) but places the endpoints in \emph{different
loss basins}: the longer run spends 50 additional low-LR epochs drifting
into a neighboring minimum whose decision boundaries disagree with the
shorter run on 13\% of test samples. Averaging softmax probabilities
across the two runs recovers $+1.1$ to $+1.5$ percentage points
\emph{beyond} what a matched 10-seed same-length deep ensemble already
delivers.

We call this trick \ours{} and validate it on nine benchmarks across EEG,
images, and text, using four architecture families. On FACED 9-class EEG
emotion classification \ours{} sets a new state of the art at
\facedacc{} balanced accuracy---$+5.3$~pp ($+8.2\%$ relative) over the
EMOD AAAI 2026 baseline, $+1.1$~pp above a 10-seed same-length deep
ensemble---and is confirmed cross-dataset on SEED-V ($\seedvgain{}$ gain).
The same single-integer change lifts CIFAR-10/100, MNIST, and
Fashion-MNIST with small-CNN, ResNet-18, and MLP backbones, and a
DistilBERT run on SST-2. Head-to-head under matched compute, \ours{} wins
on all four calibration metrics (accuracy, NLL, ECE, Brier) against SWA,
snapshot ensembles, and 10-seed deep ensembles.

A simple theory explains the entire pattern. The gain equals the
between-length disagreement fraction times the per-disagreement accuracy
recovery ($0.131 \times 0.103 = 1.35$~pp predicted, $1.50$~pp observed),
and it scales as a power law in the unused headroom $(1-p_{\mathrm{ind}})$
with exponent $\alpha=0.85$ (Pearson $r=0.75$, $p=0.019$, $R^2=0.57$).
\ours{} is therefore most valuable exactly where ensembling matters
most---on capacity-bound students near their effective ceiling.
\end{abstract}
