\begin{abstract}
\textbf{Late-layer concept directions are not free real estate for
supervision.} Once a network has absorbed a concept from task supervision
alone, adding the same concept as an auxiliary loss is at best a no-op
and at worst harmful --- a regularity we call \emph{saturation}. We
discover and quantify saturation by studying the principal axis of human
emotion (the \textbf{V-axis}, valence) across language models, human
brains, and EEG emotion classifiers.

The V-axis itself is universal: extracted from nine short emotion
stories via PCA on language-model centroids, it is shared across 14 LLMs
(560M--32B parameters across the Qwen3 family plus Llama-4, Mistral,
Phi, Pythia, BLOOM, TinyLlama, and Gemma); a single Qwen3-4B run
recovers sentiment benchmarks (SST-2 AUC $0.832$, IMDB $0.857$, Yelp
$0.938$, TweetEval $0.927$) and word-valence lexicons (Hu \& Liu,
Warriner) at $|r|\!\ge\!0.70$ zero-shot,
predicts cohort-averaged EEG of 123 human subjects watching 28 emotional
videos at $r=0.87$ ($p<10^{-9}$), and emerges spontaneously in EEG
classifiers trained on the 9-class label without any V-axis supervision
(class-PC1 $r{=}{+}0.885$, within-class residual $r{=}{+}0.738$ across
36 checkpoints). LLMs and EEG classifiers reach the same axis without
ever seeing each other's data --- direct cross-modal evidence for the
Platonic Representation Hypothesis at biological scale.

\textbf{Saturation.} We test 25 representation-alignment interventions
that inject the V-axis as auxiliary supervision (auxiliary regression,
RSA, knowledge distillation, contrastive losses, PEFT adapters,
curriculum schedules, pretrain--finetune, multi-LLM ensembles,
channel-targeted topographic losses, anger-weighted variants matched to
the analytical ceiling). Above an empirically located saturation
threshold, \emph{none help}: 16 produce statistically significant
accuracy decrements ($\Delta\text{BACC}\le-0.013$, $p<0.05$) and five
families exhibit \emph{monotonic} destruction with intervention
strength. The transition from neutral to harmful is sharp, sits in the
BACC interval $[0.62, 0.66]$ on FACED-9, and disappears at weak
baselines (CBraMod $\Delta\!\in\![-0.005, +0.005]$, all $p>0.2$).

\textbf{Mechanism and SOTA.} The within-class residual $|r|$ of an EEG
checkpoint quantitatively predicts its leave-one-out contribution to
ensemble accuracy ($r{=}{+}0.74$, $p{=}0.014$, $n{=}10$) --- the
\emph{Residual Contribution Law}. A causal Arditi-style ablation of
$v_{\mathrm{resid}}$ drops BACC well above a matched random-direction
null on every checkpoint (mean $z\approx 7.7$, all $p<0.001$). The class-mean basin is
saturated and shared; the within-class residual is seed-specific and is
the only subspace where ensemble averaging recovers signal. This yields
a new FACED 9-class SOTA: \textbf{0.6948 BACC ensemble} ($+21.5\%$
relative over CBraMod) and \textbf{0.6755 single-checkpoint}
(val-selected). \emph{What this is not}: an analytical theorem; a
within-subject claim (cohort-averaged across 123 subjects); a universal
threshold (FACED-specific). Cross-dataset replication on SEED-V is at
the class-ranking level ($r{=}{+}0.96$) plus a parallel five-claim
re-derivation (cohort $r{=}{+}0.62$, posterior dominance, cross-arch
$r{=}{+}0.601$, saturation, and causal-ablation closure with mean
$z{=}{+}23.6$).
\end{abstract}
