\section{Related Work}
\label{sec:related}

\paragraph{Brain--LM alignment.}
Toneva and Wehbe \citep{toneva2019interpreting} first demonstrated that
mid-layer transformer activations linearly predict fMRI responses to
narrative text, and that fine-tuning towards the brain target voxels
improves both the alignment and downstream NLP performance --- the
\emph{targeted-voxel} fine-tuning paradigm we cite as our closest
methodological precedent. Schrimpf et~al.\ \citep{schrimpf2021neural}
showed across 43 LMs that next-word prediction performance correlates
with brain-fit on fMRI/ECoG. Goldstein et~al.\
\citep{goldstein2022shared} extended the picture to intracranial
recordings; Caucheteux and King \citep{caucheteux2022brains} to MEG.
Padakanti and Oota \citep{padakanti2025meg} (EMNLP 2025) provide the
most direct methodological precedent for our brain-side analysis:
per-sensor ridge alignment with topographic scalp maps as the
publishable artifact. Our work differs by (i) using the EEG modality at
full 123-subject scale, (ii) studying \emph{emotion} rather than
language, (iii) showing the alignment goes both ways: LLM-derived
directions predict brain responses, and EEG models converge to the same
LLM-derived direction.

\paragraph{Affective computing on FACED.}
The FACED 9-class dataset \citep{chen2023faced} has emerged as a
standard benchmark for fine-grained EEG emotion classification.
CBraMod \citep{wang2025cbramod} (ICLR 2025) reports 0.572 BACC on the
full 9-class task with a transformer foundation model pretrained on
TUEG. EmotionKD \citep{liu2023emotionkd} (ACMMM 2023) reports 0.628 via
knowledge distillation from a labelled teacher. EMOD
\citep{chen2025emod} (AAAI 2026) reports 0.6287 with an axial
transformer backbone using V--A contrastive supervision. REVE
\citep{elouahidi2025reve} (NeurIPS 2025) sets a recent multi-task EEG
benchmark; LaBraM \citep{jiang2024labram} (ICLR 2024 Spotlight) and
EEGPT \citep{yue2024eegpt} provide additional foundation-model
references.

\paragraph{Representation alignment as auxiliary supervision.}
Aligning a target representation to an auxiliary signal --- whether
from a teacher network \citep{hinton2015distilling}, an alternative
representation \citep{tian2020crd}, a similarity matrix
\citep{kriegeskorte2008rsa,park2019rkd}, or human perceptual judgements
\citep{sundaram2024perceptual} --- is a well-studied mechanism for
inducing transferable structure. Recent work has explored cross-modal
alignment via contrastive losses \citep{radford2021clip}, RSA-style
structural matching, label smoothing relative to teacher distributions
\citep{khosla2020supcon}, and parameter-efficient adaptation
\citep{hu2022lora,liu2022ia3}. Our paper's negative results probe this
literature: when the model has already converged to the alignment
target via task supervision alone, adding any of these mechanisms
\emph{cannot} help. We test 25 such mechanisms and confirm: above the
saturation threshold, all hurt or do nothing.

\paragraph{Concept directions and probing in LLMs.}
A growing body of work demonstrates that abstract concepts often
correspond to single directions in transformer hidden states. Arditi
et~al.\ \citep{arditi2024refusal} (NeurIPS 2024) show that
\emph{refusal} in instruction-tuned LMs is mediated by a single
direction in residual-stream activations across 13 models up to 72B
parameters; ablating this direction eliminates refusal, while adding
it induces refusal. Earlier work in this style includes
\citep{li2023inferencetime,zou2023repe,kim2018tcav}. Our V-axis is
extracted analogously --- as the principal component of class-mean
activations across nine emotion-story prompts --- and exhibits the
same architecture-spanning universality.

\paragraph{The Platonic Representation Hypothesis.}
Huh et~al.\ \citep{huh2024platonic} (ICML 2024 position paper) argue
that representations learned by neural networks across modalities and
architectures converge to a shared statistical model of reality, and
operationalize this via kernel alignment between vision and language
backbones. Our work provides cross-modal evidence beyond the
vision--language pair studied there: (i) LLM $\leftrightarrow$ human
EEG (a biological neural system), and (ii) LLM $\leftrightarrow$ EEG
models (an artificial neural
network trained on biological signals), all converging to the same
one-dimensional valence axis. The convergence we measure within EEG
models trained on the same dataset is also a Platonic-style phenomenon
at the single-modality level.

\paragraph{Frontal asymmetry in emotion.}
Davidson and colleagues
\citep{davidson1992anterior,davidson2004wellbeing} proposed that
approach versus withdrawal motivation is reflected in relative
frontal-alpha asymmetry: greater left-frontal activation indexes
positive affect / approach, and right-frontal activation indexes
negative / withdrawal. The hypothesis has been replicated in many
static-stimulus paradigms and has remained the dominant
electrophysiological account of emotion in the EEG literature
\citep{coan2004frontal}. Our brain topography results show that for
\emph{video-evoked emotion} on FACED, the LLM-derived valence
direction is predominantly encoded in posterior visual cortex
($|r|_{\text{occipital}}=0.21$ vs.\ $|r|_{\text{frontal}}=0.16$), with
Davidson's frontal asymmetry replicating qualitatively but at
$\sim$10$\times$ smaller magnitude. We do not refute Davidson's
hypothesis --- the frontal effect is real --- but we add an
empirically more powerful account for video paradigms specifically.

\paragraph{Ensemble theory.}
Classical bias--variance decomposition
\citep{geman1992neural,krogh1995neural} explains uniform softmax
averaging as variance reduction across a fixed bias. Recent
mechanistic work on linear-mode connectivity
\citep{frankle2020linearmode,wortsman2022model} has shown that
checkpoints from the same training run lie in approximately convex
basins, sharing a coarse representation but differing in fine details.
Our two-tier story --- saturated class-mean basin (Tier 1) plus
seed-specific within-class residual (Tier 2) --- gives a
representational version of this picture, with the residual explicitly
identified as the \emph{V-axis} component.
