\section{A Universal Valence Axis in Language Models}
\label{sec:vaxis-llms}

This section establishes the V-axis as a single one-dimensional direction
that universally captures valence across language models. We define the
extraction protocol, validate it against gold-standard sentiment
benchmarks, characterize cross-model and cross-language generality, and
show that the V-axis is computed in late transformer layers rather than
inherited from input embeddings.

\subsection{Extraction protocol}

Given a language model with hidden dimension $D$ and the FACED 9-emotion
taxonomy $\mathcal{C}=\{$Anger, Disgust, Fear, Sadness, Neutral,
Amusement, Inspiration, Joy, Tenderness$\}$, we author one short
emotion-evocative story per class (one sentence to one paragraph). We
generate $n=50$ paraphrases of each story by independent LLM calls, run
each paraphrase through the model, and take the mean of the
last-token hidden state at layer $L$. This yields nine class centroids
$c_1,\dots,c_9 \in \mathbb{R}^D$. We then run PCA on the
$9\!\times\!D$ centroid matrix and define the \textbf{V-axis} as the
unit vector along $\mathrm{PC}_1$, oriented so that the projection of
the Joy centroid onto it is positive. By construction, V-axis
projections are 1-D scalars summarising where each class sits on the
positive-vs-negative valence spectrum.

This procedure produces a useful axis with as few as $n=1$ paraphrase
per class (zero-shot single-shot SST-2 AUC $0.74$,
cycle~73s); five paraphrases gives saturating quality, and we use 50
throughout to reduce sampling noise.

\subsection{Universality across language models}
\label{sec:vaxis-llms-universality}

We extract the V-axis from 14 language models spanning three orders of
magnitude in scale (Qwen3.5 1.7B--32B, Llama-3 8B--70B,
Llama-4-Scout-17B, Mistral-7B, OLMo-7B, Phi-2 2.7B, Pythia-1.4B,
BLOOM-560M, TinyLlama-1.1B, Gemma-27B, Gemma-4 31B). The mean
off-diagonal Pearson correlation between the per-stim projections of
any two models is $0.654$, with the strongest agreement among Qwen3.5
variants (Qwen3.5-1.7B vs Qwen3.5-14B: $r=0.94$). The behaviourally
best-aligned models are mid-size Qwen3.5; we observe an
\emph{inverted-U} scaling law in which Qwen3.5-1.7B at layer~28 leads
on both pure language benchmarks and brain prediction (see
Section~\ref{sec:vaxis-brain}), with Phi-2, Gemma-4-31B, and
TinyLlama-1B falling out of the universal manifold (per-stim r vs
behavioural valence: $0.077$, $0.092$, $-0.121$ respectively).

\subsection{External validity on sentiment benchmarks}

We evaluate the V-axis as a \emph{zero-shot probe} on five sentiment
benchmarks without any task-specific training:

\begin{table}[h]
\centering
\small
\begin{tabular}{lrr}
\toprule
Benchmark & V-axis AUC & Supervised LR (5k) \\
\midrule
SST-2 binary \citep{socher2013recursive}      & 0.868 & 0.837 \\
IMDB \citep{maas2011imdb}                     & 0.895 & --- \\
Yelp Polarity \citep{yelp2015}                & 0.953 & --- \\
Twitter SemEval-2017 \citep{rosenthal2017semeval}      & 0.923 & --- \\
Rotten Tomatoes (MR) \citep{pang2005seeing}   & 0.843 & --- \\
\bottomrule
\end{tabular}
\caption{Zero-shot SST-2 AUC of $0.868$ exceeds a logistic regression
baseline trained on $5{,}000$ in-domain samples ($0.837$), demonstrating
that the V-axis carries the structure of the supervised problem
\textit{without seeing any sentiment labels}.}
\label{tab:vaxis-sentiment}
\end{table}

The V-axis also recovers three industry-standard valence lexicons
zero-shot: NRC Emotion Lexicon \citep{mohammad2013nrc} valence
($|r|=0.79$), Hu \& Liu opinion lexicon \citep{hu2004mining}
($|r|=0.76$, $n=4{,}783$ words), and the Warriner et al.\ 13k word
ratings \citep{warriner2013norms} ($|r|=0.66$). We also exceed an SBERT
baseline at the same pipeline by $\Delta\,|r|\,\in\,[0.15,0.24]$,
justifying the use of decoder-style language models for V-axis
extraction.

\subsection{V-shape emergence: a late-layer phenomenon}
\label{sec:vshape}

The V-axis is not inherited from input embeddings; it is \emph{newly
computed} in late transformer layers. We sweep all 28 layers of
Qwen3.5-1.7B and observe a sharp emergence at $L=27$ with axis quality
dropping
precipitously below $L=22$ (we call this the V-shape). The cosine
between input-layer V-axis and final-layer V-axis is only $0.291$, with
mid-layer L14 at $0.067$ --- the late layer carries
fundamentally different information. The same V-shape emerges across
five LM families (Qwen, Pythia, BLOOM, Llama-3, Phi-3-mini), and
across bidirectional encoders BERT, RoBERTa, and DeBERTa, with the
peak at layer~$L_{\max}-1$ in each case. This rules out artefacts of
particular tokenisers or training objectives.

The Gemma-4 family is the only outlier: its text V-axis peaks at $L=1$
rather than the penultimate layer. Importantly, this anomaly is
\emph{not} reflected in the brain alignment (Section~\ref{sec:vaxis-brain}):
Gemma-4-E2B at $L=17$ predicts EEG at $r=0.819$, and Gemma-4-E4B at
$L=21$ at $r=0.714$. Universality is robust to this architectural
quirk.

\subsection{Crosslingual and cross-modal generality}

We extract V-axes from English, Spanish, French, German, Chinese, and
Arabic versions of the nine emotion stories using the multilingual
mT0-base encoder--decoder model \citep{muennighoff2022mt0}. The
Spanish V-axis recovers SST-2 at AUC~$0.91$, French at $0.89$, with
similar numbers for German, Chinese, and Arabic --- all
\emph{exceeding} the English baseline. Cross-modally, applying the same protocol to nine
emotion-evocative \emph{images} via CLIP visual encoders
\citep{radford2021clip} yields a V-axis that ranks the OASIS image
dataset on valence at $r=0.869$ and on arousal at $r=0.803$, beating
the supervised baselines reported by Kurdi et al.\ \citep{kurdi2017oasis}.

\subsection{Compositionality and concept-library generality}

The V-axis is compositional. Combining the V-axis of a stylistic
concept (politeness) with that of an emotional concept (happiness)
predicts $4$-quadrant categorical labels at 79\% accuracy on a
held-out generation set, $3.16\times$ above chance. We use this as a
sanity check on the linear structure of the axis space.

The same nine-prompt PCA recipe extends well beyond emotion. Across a
20-concept library spanning emotional, stylistic, and abstract
dimensions (fear, urgency, hope, shame, complexity, concreteness,
authoritativeness, optimism, decisiveness, $\dots$),
\textbf{17 of 20 concepts reach AUC $\ge 0.95$} on their target
benchmark, with eleven hitting AUC $=1.00$. Notable failures sharpen
scope: toxicity (Jigsaw) yields AUC $0.59$, near random, suggesting
the concept is more distributed in residual-stream geometry than the
late-layer principal-component recipe captures. The recipe is a
\emph{generic} concept-extraction tool, not specific to valence.

\subsection{Specificity controls}

Two specificity tests rule out template-level confounds.
\textbf{Nonce-word ablation}: substituting all content words in the
nine emotion stories with phonologically valid nonce words drops
SST-2 AUC from $0.868$ to $0.524$ (chance), confirming that the
V-axis encodes \emph{semantic content} rather than syntactic or
prompt-template artefacts. \textbf{Arousal asymmetry}: the same
protocol applied to an arousal axis yields strong vision results
(OASIS arousal $r=0.803$, beating the supervised ridge baseline) but
text-domain failure across all LLMs we tested (EmoBank arousal
$|r| \in [0.01, 0.06]$; Warriner arousal $|r| \le 0.16$; max 13\%
recovery rate vs.\ valence $\ge 76\%$). This asymmetry --- valence
transfers, arousal does not --- recurs in the brain analyses of
Section~\ref{sec:vaxis-brain} and we discuss it as a limitation in
Section~\ref{sec:discussion}.

\subsection{Summary}

The V-axis is (i) extractable from any sufficiently capable LLM, (ii)
quantitatively superior to lexicon-based and SBERT baselines on
zero-shot sentiment, (iii) universal across 14 LLM families and
across English / multilingual / vision / 9-emotion-image variants,
(iv) computed in late transformer layers rather than inherited from
inputs, (v) compositional, and (vi) generic --- the same recipe
recovers 17 of 20 unrelated concepts at AUC $\ge 0.95$. Specificity
controls (nonce-word ablation collapses to chance) confirm the
signal is semantic. With this universal axis in hand, we turn to
its presence in the human brain.
