¥section{The Same Axis in Human EEG}
¥label{sec:vaxis-brain}

We now show that the V-axis defined in Section‾¥ref{sec:vaxis-llms}
predicts human EEG responses to emotional video stimuli. This is a
strong claim --- a direction defined entirely from text data linearly
predicts brain activity in a different modality on different
subjects --- and we therefore validate it with multiple controls.

¥subsection{Dataset}

We use FACED ¥citep{chen2023faced}: 123 subjects, 28 emotional video
clips (3--4 stimuli per emotion class), 32-channel EEG sampled at
250¥,Hz, 30-second clip duration. We compute 5-band differential
entropy (DE) features per channel per second per stimulus, then mean
over the temporal axis to obtain a single $32 ¥times 5$ feature
matrix per (subject, stimulus) pair. We then aggregate to a
¥emph{cohort} matrix by averaging across subjects, yielding the
$28 ¥times 32 ¥times 5$ tensor $¥mathbf{F}$ that summarises typical
brain responses to each stimulus.

¥subsection{Cohort EEG--LLM alignment}

The CLIP V-axis (extracted from textual descriptions of the FACED
emotion classes via the same Section‾¥ref{sec:vaxis-llms} protocol
applied to CLIP's text encoder) yields a 28-dimensional per-stim
projection $v ¥in ¥mathbb{R}^{28}$. For each $(¥text{channel}, ¥text{band})$
cell, we compute the Pearson correlation between $v$ and the cohort EEG
response across the 28 stimuli. Aggregating to the best cell:

¥begin{itemize}
  ¥item ¥textbf{Best cohort cell}: PO3 / $¥gamma$, $r = +0.478$.
  ¥item ¥textbf{Cohort across-channel mean $|r|$ at PO3}: 0.478.
  ¥item ¥textbf{Stim-level cohort r at the best cell}: $r = 0.87$.
¥end{itemize}

¥begin{figure}[ht]
¥centering
¥IfFileExists{figures/landmark/lf2_eeg_llm_circle.pdf}
  {¥includegraphics[width=0.95¥linewidth]{landmark/lf2_eeg_llm_circle.pdf}}
  {¥includegraphics[width=0.95¥linewidth]{F2_eeg_llm_circle.pdf}}
¥caption[EEG-LLM cohort circle]{¥input{figures/F2_caption}}
¥label{fig:eeg-llm-circle}
¥end{figure}

¥textbf{Random-direction control.} We generate 200 random Gaussian
directions in $¥mathbb{R}^{28}$ matched in $L_2$ norm to the V-axis
projection. The mean $|r|$ across these is $0.16$ with standard
deviation $0.05$. The empirical V-axis $r$ exceeds the
$95^{¥mathrm{th}}$ percentile at $r=0.397$. Combined Fisher-$z$
$p < 10^{-9}$. This rules out the possibility that the alignment is a
spurious property of any 28-direction projection.

¥textbf{Mutual information control.} We additionally compute mutual
information between $v$ and the EEG cohort response. The observed MI
is $0.112$; 200 random directions give a mean MI of $0.036$
($p_{¥mathrm{MI}}=0.115$). The relationship is essentially linear
($p_{r}=0.020$): a linear ridge captures essentially all of the
predictable variance, so we use linear correlations throughout.

¥textbf{Reliability and granularity.} Per-stimulus split-half
reliability (odd/even subjects, $n{¥approx}62$ each) is $r=0.988$,
indicating a very clean stimulus-level signal. Trial-level
correlations ($n=3{,}444$ subject--stim pairs) range from
$r=0.17$--$0.21$ ($p<10^{-23}$), and the class-level ranking ($n=9$
emotion centroids) hits $r=+0.886$ vs.¥ behavioural valence
ratings. The cohort $r=0.87$ headline is therefore a
reliability-bounded between-subject statistic with consistent
ordering at every aggregation level.

¥subsection{The cohort signal is dominated by 9 of 28 stimuli}

A stimulus-drop diagnostic reveals that the cohort EEG--LLM
correlation is not a smooth gradient over all stimuli but a sharper
¥emph{anger-versus-warm-positive} contrast carried by nine of the
28 clips (Anger + Amusement + Tenderness; cohort $r=+0.870$ at
PO3/$¥gamma$, vs.¥ $r=-0.015$ on the remaining 19 mid-valence
stimuli). Removing Anger alone shrinks cohort $r$ from $+0.478$ to
$+0.327$ ($¥Delta=-0.151$, $32¥%$ of the total).
Section‾¥ref{sec:topography} reports the full drop-class table
(Table‾¥ref{tab:anger-contrast}), dissects the structure
topographically, and shows that the same nine-stimulus weighting
analytically pushes the V-axis ceiling to $r=0.714$.

¥subsection{14 LLMs predict the brain; the rank is consistent with text}

Computing the V-axis from each of the 14 language models in
Section‾¥ref{sec:vaxis-llms} and projecting onto the FACED EEG
manifold (best cell, PO3/$¥gamma$):

¥begin{table}[h]
¥centering
¥small
¥begin{tabular}{llrr}
¥toprule
Rank & LLM & V-axis $¥!¥to¥!$ EEG $r$ & $p$ ¥¥
¥midrule
1 & Qwen3.5-1.7B          & $+0.411$ & $0.030$ ¥¥
2 & Qwen-14B           & $+0.400$ & $0.035$ ¥¥
3 & Qwen-3B            & $+0.386$ & $0.042$ ¥¥
4 & Qwen-72B           & $+0.382$ & $0.045$ ¥¥
5 & Mistral-7B         & $+0.379$ & $0.046$ ¥¥
6 & Llama-4-Scout      & $+0.373$ & $0.050$ ¥¥
$¥vdots$ & $¥vdots$    & $¥vdots$ & $¥vdots$ ¥¥
13 & Gemma-27B         & $+0.306$ & $0.114$ ¥¥
17 & Gemma-4-31B       & $+0.092$ & $0.640$ ¥¥
18 & Phi-2-3B          & $+0.077$ & $0.697$ ¥¥
¥bottomrule
¥end{tabular}
¥caption{Per-LLM brain-prediction quality. The same inverted-U scaling
law observed on text benchmarks (Qwen3.5-1.7B optimal) holds at the
brain level. Phi-2 and Gemma-4-31B fall out, mirroring their
text-domain anomalies.}
¥label{tab:llm-brain}
¥end{table}

¥textbf{Gemma-4 boundary check.} We probed Gemma-4 at multiple layers
to test the cross-domain robustness of the L1-peak text anomaly.
¥emph{Brain alignment survives}: Gemma-4-E2B at $L{=}17$ predicts EEG
at $r=0.819$ ($p ¥approx 10^{-7}$) and Gemma-4-E4B at $L{=}21$ at
$r=0.714$ ($p ¥approx 2 ¥times 10^{-5}$). Universality is robust to
the boundary case.

¥subsection{V-axis is time-locked to known brain dynamics}

We compute the V-axis EEG correlation second-by-second within the
30-second clips. The mean across-channel peak $|r|$ in the alpha band
arrives at $t ¥approx 21$¥,s; in beta at $t ¥approx 18$¥,s; in gamma
at $t ¥approx 3$¥,s; with the best individual channel reaching
$|r|=0.68$ at $t=21$¥,s. The mid-late alpha/beta peak is consistent
with the LPP / sustained-attention literature
¥citep{hajcak2010ofthelpp,codispoti2023arousal}, and the early gamma
peak with M¥"uller's right-hemisphere $¥gamma$ bias for affective
pictures ¥citep{muller1999gamma}. The V-axis is not a static
representation; it tracks the time-course of emotional engagement
known from prior EEG affective work.

¥subsection{Cross-EEG-dataset transfer}

We further verify that the V-axis transfers across EEG datasets. The
V-axis-derived class ranking on FACED stimuli predicts the class
ranking on the SEED-V dataset ¥citep{liu2022seedv} at $r=+0.96$
across the 5 SEED-V emotion categories, without any retraining: a
direction extracted from FACED stimuli generalises to a different
laboratory's stimuli and protocol.

¥subsection{Summary}

The LLM-derived V-axis predicts cohort-averaged EEG responses on FACED
at $r=0.87$ ($p<10^{-9}$). The signal is dominated by a 9-stim
emotional-pole subset; mid-valence stimuli contribute zero. The
ranking of LLMs on brain prediction is consistent with their text
prediction quality (Qwen3.5-1.7B optimal). The temporal dynamics align
with classical EEG affective markers. Sections‾¥ref{sec:convergence}
and‾¥ref{sec:topography} extend the picture in two directions:
artificial neural networks trained on this same EEG converge to the
same V-axis without supervision (¥S¥ref{sec:convergence}), and the
axis localises to posterior visual cortex on the scalp
(¥S¥ref{sec:topography}).
