\section{The Saturation Regularity: V-axis as Supervision}
\label{sec:saturation}

Can explicit V-axis supervision improve EEG-emotion classifiers?
Across 25 distinct intervention families spanning seven years of
representation-alignment literature, the answer is uniformly \textbf{no}
above a saturation threshold.

\begin{quote}
\textbf{Saturation Regularity.} \emph{Let $f_\theta$ be a classifier
whose penultimate features encode a target concept direction $v$ at
strength $\rho(\theta) := |\langle \mathrm{PC}_1(H_{\mathrm{class\text{-}mean}}),
v\rangle|$. There exists a saturation threshold $\rho^*$ such that for
$\rho(\theta) > \rho^*$, supplementing the task loss with any auxiliary
loss $\mathcal{L}_v(H, v)$ at $\lambda > 0$ produces
$\mathbb{E}[\Delta\mathrm{BACC}] \leq 0$ across the families tested.}
\end{quote}

This is an \emph{empirical} statement, not a theorem. We support it
with: (i) 16 / 25 interventions producing statistically significant
accuracy decrements at the strong recipe ($p<0.05$, paired across 5
seeds; 14 sig.\ negative + 2 catastrophic); (ii) five intervention
families showing \emph{monotonic} destruction as $\lambda$ increases
(no positive setpoint exists); (iii) a unidirectional saturation
transition from neutral at weak baselines (CBraMod, EMOD-vanilla, all
$p>0.2$) to significantly harmful at strong recipes
($\Delta\!\in\![-0.024, -0.015]$, $p<0.05$); and (iv) a direct
mechanism check showing V-axis supervision moves class-PC1 by
$+0.01$--$+0.36$ but moves the within-class residual by only
$\sim 10^{-7}$ --- numerical zero. For FACED-9 the empirical threshold
sits in $\rho^* \in [0.62, 0.66]$.

\paragraph{Twenty-five interventions, one verdict.}
The families: auxiliary regression (EMODSTYLE), topographic masking
(frontal, occipital, topo-optimal, Procrustes), Davidson FAA
\citep{davidson1992anterior}, \textbf{anger-weighted} matching the
analytical 9-stim ceiling, RSA \citep{kriegeskorte2008rsa}, Distance-CE
\citep{park2019rkd}, Multi-V, soft-target KD \citep{hinton2015distilling},
PEFT \citep{hu2022lora,liu2022ia3}, pretrain-finetune, curriculum,
init-priors, cross-dataset, data-fraction scaling, uncertainty-weighted
multi-task \citep{kendall2018uncertainty}, InfoNCE
\citep{vandenoord2018cpc}, ensemble extension. Largest significant
decrements: RSA $\lambda{=}5.0$ $\Delta=-0.093$ ($p<10^{-4}$);
Distance-CE $\tau{=}5.0$ $-0.397$ ($p<10^{-4}$, chance); pretrain-FT
frozen $-0.401$ ($p<10^{-4}$, catastrophic); uncertainty multi-task
$-0.067$ ($p<10^{-3}$); RSA $\lambda{=}1.0$ $-0.057$ ($p<10^{-3}$);
anger-weighted $\lambda{=}0.5$ $-0.054$ ($p=0.0003$); frontal-mask
$\lambda{=}0.5$ $-0.052$ ($p=0.0015$); SCALING $-0.045$ ($p<0.01$);
FAA $\lambda{=}0.5$ $-0.044$ ($p=0.006$); topo-optimal $\lambda{=}0.1$
$-0.013$ ($p=0.039$). Full 25-row table in
Appendix~\ref{app:full-vaxis-table}.

\paragraph{Monotonic destruction.}
Five intervention families exhibit monotonic accuracy destruction as
the V-axis weight is increased, the cleanest form of saturation
evidence: RSA $\lambda \in \{0, 0.1, 0.5, 1.0, 5.0\}$ gives $0.624 \to
0.610 \to 0.585 \to 0.566 \to 0.531$; Distance-CE $\tau \in \{0.5, 1.0,
2.0, 5.0\}$ gives $0.516 \to 0.476 \to 0.395 \to 0.227$ (chance);
Multi-V $\lambda \in \{0.05, 0.1, 0.5\}$: $0.622 \to 0.603 \to 0.581$;
EEG-AUX MSE $\lambda \in \{0.1, 0.5\}$: $-0.016 \to -0.043$; SCALING
$\lambda{=}0.5$ at $f \in \{0.1, 0.25, 0.5, 1.0\}$: $-0.002 \to -0.029
\to -0.032 \to -0.045$. Stronger V-axis supervision is strictly worse
--- there is no positive setpoint to find.

\paragraph{Locating the saturation transition.}
V-axis supervision is within seed noise at weak baselines (CBraMod
$0.572$+Topo $\Delta{=}-0.005$, $+$EMODSTYLE $+0.005$, EMOD-$d{=}3$
vanilla+Topo $+0.007$; all n.s.\ at $n{=}4$--$5$) and produces a
significant accuracy decrement at the strong recipe (EMOD-$d{=}6$
full+Topo $\Delta{=}-0.015$, $+$EMODSTYLE $\Delta{=}-0.024$; both
$p<0.05$); the transition is unidirectional in BACC $[0.62, 0.66]$
(saturation cliff: Figure~\ref{fig:saturation-cliff},
Appendix~\ref{app:saturation-extras}). Anger-weighting analytically
maximises the V-axis ceiling ($r{=}0.714$ vs.\ $0.478$) yet is the
worst single intervention ($\Delta{=}-0.054$, $p=0.0003$); even Path~B
(adding 15--25 V-axis-trained ckpts to our 10-vanilla pool) yields a
negative ensemble delta ($\Delta\!\in\![-0.0193, -0.0145]$,
$p<0.001$). The 25 interventions
reduce to one principle:
\emph{saturated representation = unhelpful supervision target}.
