\section{State of the Art via Principled Ensembling}
\label{sec:sota}

\begin{table}[h]
\centering
\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{minipage}{0.43\linewidth}
\centering
\begin{tabular}{@{}lc@{}}
\toprule
Method & BACC \\
\midrule
CBraMod \citep{wang2025cbramod}    & $0.572$  \\
EmotionKD \citep{liu2023emotionkd} & $0.628$  \\
EMOD \citep{chen2025emod}          & $0.6287$ \\
\midrule
EMOD + aug + KD + d6 (1 seed)      & $0.6581$ \\
\textbf{Best single (val-sel)}     & $\mathbf{0.6755}$ \\
\textbf{10-ckpt ensemble (SOTA)}   & $\mathbf{0.6948}$ \\
\bottomrule
\end{tabular}
\subcaption*{(a) Headline. $+0.123$ abs.\ ($+21.5\%$) over CBraMod.}
\end{minipage}\hfill
\begin{minipage}{0.55\linewidth}
\centering
\begin{tabular}{@{}lcc@{}}
\toprule
Recipe step & BACC & $\Delta$ \\
\midrule
EMOD ($d{=}3$, $f{=}128$) replication & $0.6194$ & --- \\
$+$ augmentation                      & $0.6343$ & $+0.015$ \\
$+$ aug $+$ KD                        & $0.6439$ & $+0.025^{*}$ \\
$+$ depth doubling ($d{=}6$)          & $0.6581$ & $+0.014$ \\
$+$ longer training ($e{=}150$)       & $0.6581$ & $0.000$ \\
\midrule
5-seed e100 ensemble                  & $0.6798$ & $+0.022$ \\
\textbf{10-ckpt e100$+$e150 (SOTA)}   & $\mathbf{0.6948}$ & $\mathbf{+0.037}$ \\
\bottomrule
\end{tabular}
\subcaption*{(b) Recipe cascade ($^{*}$super-add., $p{=}3\!\times\!10^{-4}$).}
\end{minipage}
\caption{New FACED 9-class SOTA. The \texttt{rand9} ablation replaces
LLM-derived KD prototypes with random orthonormal 9-D directions and
costs $\le 0.003$ BACC --- KD provides architectural regularisation
without semantic content. Visual cascade (LF7) in
Appendix~\ref{app:ensemble-extras}.}
\label{tab:sota}
\end{table}

The cascade starts from a faithful EMOD replication at $0.6194$ and
walks up in six steps. Augmentation---temporal jitter, channel
dropout, amplitude scaling, and Gaussian noise applied with
$p{=}0.6$---contributes the largest single jump ($+0.0149$),
reflecting the limited stimulus count per subject. Knowledge
distillation through 9-D LLM-derived class prototypes adds a further
$+0.0096$; a \texttt{rand9} ablation that swaps the prototypes for
random orthonormal directions costs at most $0.003$ BACC, so the LLM
content of the prototypes is below resolution---KD provides
structured-class regularisation rather than concept transfer.
Depth-doubling to $d{=}6$ adds another $+0.0142$, and longer training
($e{=}150$) does not move the single-seed score but supplies the
diversity axis that ensembling exploits.

\paragraph{Two-tier ensemble theory.}
The $10$ $d{=}6$ checkpoints share an \emph{identical} class-PC1 basin
($|r|\in[0.60, 0.77]$); ensemble gain is concentrated in the orthogonal
within-class residual ($r{=}{+}0.743$, $p{=}0.014$). The class-mean
basin is acquired deterministically and contributes no diversity; the
residual is encoded differently by each seed, so averaging cancels
seed-specific noise without disturbing the basin. Mixing $e{=}100$
and $e{=}150$ ties at $0.6581$ single-seed but provides the most
ensemble-relevant diversity axis: the mixed $10$-checkpoint pool
reaches $0.6948$ versus $0.6798$ for $e{=}100$ alone. The mechanism
generalises (Appendix~\ref{app:ensemble-extras}): $\Delta$BACC is
monotonic in headroom (near zero on MNIST), mega-ensembles plateau at
ten checkpoints ($\le 0.6948$), and cross-architecture mixing adds no
diversity---intra-architecture, training-length-driven diversity is
what the within-class residual mechanism predicts.
