\section{Discussion}
\label{sec:discussion}

\subsection{Three sides of the same triangle}

We have shown that a single one-dimensional axis --- defined entirely
from text via principal-component analysis of nine emotion-evocative
stories --- predicts:
\begin{enumerate}
\item Five sentiment benchmarks at zero-shot AUC up to $0.953$
(Yelp), \emph{exceeding} a logistic-regression baseline trained on
$5{,}000$ in-domain samples on SST-2 (Section~\ref{sec:vaxis-llms}).
\item Cohort EEG responses to 28 emotional video clips at $r=0.87$
($p<10^{-9}$), with the signal localised to posterior visual cortex
and consistent with the LPP / sustained-attention literature
(Sections~\ref{sec:vaxis-brain}, \ref{sec:topography}).
\item EEG-emotion classifier accuracy, with cross-architecture
correlation $r=+0.885$ between V-axis encoding strength and BACC, and
within-class V-axis residual correlation $r=+0.74$ ($p=0.014$) with
ensemble contribution (Section~\ref{sec:convergence}).
\end{enumerate}

This is cross-modal evidence beyond the vision--language pair studied
by Huh et~al.\ \citep{huh2024platonic}: \emph{the same one-dimensional
direction emerges in language models trained on text, in human brains
exposed to video, and in EEG models trained on those brains}. We read
this as direct empirical support for the Platonic Representation
Hypothesis, extended to a biological neural system.

\subsection{The saturation regularity and representation alignment}

The 25-intervention saturation result
(Section~\ref{sec:saturation}) carries a strong negative for the
representation-alignment literature: \emph{external alignment losses
cannot help once the model has absorbed the alignment target from
task supervision alone}. This refines the conventional reading of
distillation, RSA, contrastive losses, and PEFT alignment: those
mechanisms work in the under-trained regime, where any well-structured
target is informative, and break down sharply at saturation.

The transition we observe is sharp and recipe-dependent. On FACED, it
sits in the BACC interval $[0.62, 0.66]$, which is approximately
where every recently-published method (CBraMod, EmotionKD, EMOD)
already lives. We do \emph{not} claim that representation alignment is universally
useless --- weak baselines below the threshold are at worst unaffected
by V-axis supervision (CBraMod 5-seed $\Delta \in [-0.005, +0.005]$,
EMOD-vanilla $\Delta=+0.007$, all within seed noise). The clean
finding is the \emph{transition}: V-axis goes from neutral at weak
baselines to statistically significantly harmful at strong recipes.
The saturation threshold for late-layer concept directions is a
real, measurable, and surprisingly low watermark.

\subsection{Two-tier ensembling and the source of practical gains}

The two-tier picture (Section~\ref{sec:two-tier}) gives a precise
mechanistic account of where ensemble gains come from on saturated
networks. Tier 1 (the class-mean V-axis basin) is shared across seeds
and contributes nothing to ensemble diversity; Tier 2 (the
within-class V-axis residual) is seed-specific and is the source of
the entire $+3.7$ percentage points we measure. This generalises
classical bias--variance decomposition
\citep{geman1992neural,krogh1995neural} to a representation-level
statement: ensemble gain on a saturated network is variance reduction
\emph{within the orthogonal complement of the converged subspace}.

The same mechanism explains the well-known plateau of
"mega-ensembles" --- once 5--10 seeds saturate the residual
variance reduction, additional checkpoints add correlated noise. We
observe this directly: 15-, 20-, 25-checkpoint pools all match or
underperform our 10-checkpoint SOTA.

\subsection{Implications for affective neuroscience}

The brain-side findings are stronger than we initially anticipated.
For video-evoked emotion on FACED, the V-axis is encoded
predominantly in posterior visual cortex
($|r|_{\mathrm{occipital}}=0.21$ vs.\ $|r|_{\mathrm{frontal}}=0.16$),
and Davidson's frontal-alpha asymmetry replicates in direction at
smaller magnitude than the posterior encoding. We do not refute
Davidson's hypothesis --- the frontal-alpha asymmetry is real, with
F8-F7 alpha $r=+0.0155$ and Fp2-Fp1 alpha $r=+0.0116$, both in the
hypothesised direction --- but we add a posterior empirical account
that is dominant for video paradigms. This complements
the \emph{static-stimulus} literature
\citep{davidson1992anterior,coan2004frontal} where Davidson's
asymmetry was primarily characterised, and we read the magnitude
difference as a paradigm effect (dynamic visual stimuli engaging
posterior visual cortex) rather than a contradiction of within-subject
FAA effects.

The 9-stimulus emotional-pole structure (Section~\ref{sec:topography})
sharpens the V-axis claim from "smooth gradient over all emotions" to
"anger-versus-warm-positive contrast across nine clips". The Simpson's-paradox
dynamic --- cohort $r=0.48$, per-subject mean $r=-0.06$, oracle
$|r|=0.62$ --- locates the population effect at the
\emph{between-subject} level, with substantial inter-subject
heterogeneity in how the V-axis is topographically encoded. This is
consistent with Sani et~al.\ \citep{sani2024crosssubject} and
suggests that within-subject EEG-emotion modelling will benefit from
per-subject channel selection, while between-subject claims (such as
ours) are best made cohort-side.

\subsection{Honest negatives}

We document a number of negatives that are diagnostic rather than
disqualifying. Each shapes the scope of one of our positive claims.

\paragraph{Theta-gamma coupling does not mediate the V-axis.}
The Tort modulation index across channels gives $\rho(\mathrm{PAC},
\mathrm{V\text{-}axis\ }r) = +0.082$, $p=0.667$. Phase-amplitude
coupling \citep{canolty2010cfc} is the most popular
\emph{electrophysiological} candidate mechanism for emotional
cognitive control, and it is fully decoupled from the V-axis
encoding in our data. The V-axis is a cohort-DE-band linear
projection phenomenon; it is not phase-amplitude mediated.

\paragraph{Mutual information control is not significant.}
$p_{\mathrm{MI}}=0.115$ vs.\ linear $p_r=0.020$ on the same
random-direction null. The V-axis--EEG relationship is essentially
linear: a non-linear similarity estimator does not detect signal
above what linear regression captures. This licences our use of
ridge regression as the right-sized analysis tool but rules out
richer non-linear claims (e.g.\ a sigmoidal saturation in brain
encoding, or band-band non-linear interactions). Future work
seeking non-linear V-axis structure must clear this MI bar.

\paragraph{Random-direction null places class-PC1 at 93rd percentile.}
The cross-architecture class-PC1 r=+0.885 sits at the $93^{\mathrm{rd}}$
percentile of a 100-direction matched-norm Gaussian null
($p_{\mathrm{one}}=0.079$). The 9-class PC1 subspace is intrinsically
low-dimensional, which gives a wide null distribution; an honest
reading is that class-PC1 alignment is \emph{above-random but not
uniquely so}. Our statistically robust claim is the within-class
residual: $r=+0.74$, $p=0.014$ between residual $|r|$ and
ensemble contribution. We foreground that as the primary claim and
treat class-PC1 as suggestive supporting evidence.

\paragraph{Arousal asymmetry: vision yes, text and brain no.}
An arousal axis extracted via the same protocol predicts OASIS
image-arousal at $r=0.803$ (beating supervised ridge $0.789$).
However, the same axis fails on text (EmoBank A
$|r| \in [0.01, 0.06]$; Warriner A $|r| \le 0.16$; max 13\%
recovery rate vs.\ valence's $\ge 76\%$) and is substantially
weaker on FACED EEG ($r \in [0.18, 0.41]$ across LLMs). Cross-language
A-axes are essentially orthogonal. The asymmetry is consistent with
the cross-modal arousal observation of \citep{kurdi2017oasis}, where
arousal is more strongly modality-conditioned than valence. We
explicitly limit V-axis brain claims to valence and discuss arousal
extension as future work.

\paragraph{LLM-content of KD is irrelevant at this scale.}
Replacing the 9-D LLM-derived class prototypes with random
orthonormal 9-D directions changes BACC by $\le 0.003$
(\texttt{rand9} ablation, $0.6467$ vs.\ LLM-9D $0.6464$). At strong
baselines, KD provides architectural regularisation; its semantic
content is below our resolution. This is itself a saturation
signature: by the time KD is applied, the network has already
absorbed the class structure from the labels, and the soft-target
geometry is a rotational invariant in the loss landscape.
Importantly this does \emph{not} contradict the V-axis story: it
specifically says that injecting V-axis information \emph{through
KD} fails for the same reason all 25 V-axis interventions fail at
the strong baseline (Section~\ref{sec:saturation}).

\paragraph{Per-subject V-axis adaptation does not transfer.}
The per-subject best-channel oracle reaches $|r|=0.62$ (a $+0.39$
absolute headroom over the cohort-fixed top-8 of $|r|=0.02$). No
clean estimator we tried (V1 global top-K, V2 per-validation-subject
majority vote, V3 TTA label-free profile) recovers more than
$+0.001$ over the cohort top-K. This is documented in
Appendix~\ref{app:negatives}: the per-subject signal is real but
not transferable through the simple selection rules we tested. A
subject-conditioned channel-attention head trained on a small
calibration set per subject is the natural follow-up
(Section~\ref{sec:discussion}, future work).

\paragraph{Cross-architecture ensembling does not add diversity.}
Mixing $d \in \{4, 6, 8, 10\}$ EMOD checkpoints gives within-arch
x-seed BACC $0.6798$ vs.\ cross-arch $0.6791$ (n.s.). The diversity
that helps is intra-architecture, training-length-driven (mixing
$e{=}100$ and $e{=}150$ at fixed $d{=}6$). We use this to argue that
diversity in the within-class V-axis residual subspace is the lever,
not depth or capacity diversity per se.

\subsection{Limitations}

\textbf{One emotion benchmark, one EEG dataset.} FACED 9-class is the
strongest available 9-emotion benchmark, but the V-axis story would
benefit from replication on richer affective datasets (e.g., DEAP,
DREAMER, MAHNOB-HCI). We provide a partial cross-dataset transfer
result (FACED $\to$ SEED-V class ranking $r=+0.96$) but the strongest
claim is FACED-internal.

\textbf{Cohort-side analysis.} Our brain alignment claims are made on
123-subject \emph{cohort-averaged} EEG, not within-subject. The
Simpson's-paradox dynamic of Section~\ref{sec:topography} is a
limitation of this design. Within-subject decoding requires
per-subject channel adaptation that we tested but did not find
to transfer (the per-subject best-channel oracle reaches $|r|=0.62$
but no clean estimator approaches this).

\textbf{Saturation threshold is dataset-specific.} The
$[0.62, 0.66]$ BACC window we identify is the FACED-9 saturation
range. Other tasks will have different thresholds, set by the
intrinsic difficulty of the labels and the capacity of the encoder.
The qualitative shape of the saturation transition --- "helps weak,
hurts strong" --- generalises in our cross-dataset checks
(\S\ref{sec:saturation}); the precise location does not.

\textbf{Negative results require the strong-baseline interpretation.}
Twenty-five interventions failing at our recipe is consistent with
saturation but also consistent with deficiencies of those particular
intervention designs. We addressed this by sweeping each family
across $\lambda$ values, including monotonic-destruction families
that rule out "we used the wrong scale" as the explanation.

\subsection{Future work}

\textbf{Per-subject V-axis adaptation.} The per-subject best-channel
oracle ($|r|=0.62$) suggests substantial untapped within-subject
signal. A subject-conditioned channel-attention head trained on a
small calibration set per subject would close the cohort-vs-individual
gap in a paper-grade way.

\textbf{Other concept directions.} The saturation regularity is stated
for the V-axis; its generality across other late-layer concept
directions \citep{arditi2024refusal} is open. Replicating the
finding for arousal, dominance, formality, or refusal directions on
their respective benchmarks is the natural next step.

\textbf{Saturation as a model-selection signal.} The within-class
residual $|r|$ is computable from any checkpoint and predicts
ensemble contribution. Using it as an online model-selection signal
during training (rather than as a post-hoc diagnostic) is a concrete
engineering follow-up.

\subsection{Conclusion}

A single one-dimensional valence axis, extracted from nine short
emotion-evocative stories, predicts language-model behaviour, EEG
responses, and EEG classifier accuracy with quantitatively consistent
strength. We frame this as cross-modal evidence for the Platonic
Representation Hypothesis at biological scale, document a saturation
regularity that bounds when this universal axis can serve as
supervision, and operationalise the residual variance into a new
FACED SOTA. The brain-side picture --- posterior visual cortex,
direction-consistent but smaller-magnitude frontal asymmetry,
9-stimulus emotional-pole contrast,
Simpson's-paradox between subjects --- positions the V-axis claim
firmly within the affective-EEG literature while substantially
extending it.
