\begin{abstract}
Can a useful sentiment classifier be extracted from a language model with
\emph{zero} human labels, nine generated stories per class, and three lines
of linear algebra? We show that the answer is yes, and that the resulting
direction is startlingly universal. We introduce \ours: generate nine short
stories for each of nine emotion categories (81 LLM-generated paragraphs,
no polarity labels), mean-pool the final-layer residual stream, and take
the first principal component of the nine category centroids as a
\emph{valence axis}. Projecting held-out text onto this single direction
gives AUC \sstauc{} on SST-2, \imdbauc{} on IMDB, \yelpauc{} on Yelp
Polarity, and \twauc{} on Twitter SemEval-17, correlates \huliur{} with
the Hu~\& Liu 6{,}789-word lexicon and \warrinerr{} with the Warriner
continuous valence norms --- all from the \emph{same} 81 English stories
and no supervised training. Applied zero-shot, the \emph{English} axis
transfers to held-out stories in Spanish (AUC $0.949$), French ($0.991$),
German ($0.972$), and Chinese (\emph{$1.000$}), without translation or
re-extraction. A supervised logistic regression on the same hidden
features needs $N\!\approx\!100$ SST-2 labels to match our zero-label
AUC, and the axis is stable across model scale (Qwen2.5-1.5B and 7B both
reach $|r|\!=\!0.494$ on EmoBank) and family (Qwen, Pythia, BLOOM all
produce functionally aligned per-story rankings at mean cosine $0.961$).
The recipe \emph{also} beats the standard contrastive mean-difference
extraction of Zou et al.\ and Arditi et al.\ on matched features by
$+0.115$ AUC ($0.868$ vs.\ $0.753$), without requiring the operator to
supply a positive/negative split. A layer-wise sweep reveals a sharp
``V-shape'': valence is present at layer 1, fades through the
context-integration middle, and \emph{re-crystallizes} at the final two
transformer blocks --- a pattern universal across three causal families
and absent in bidirectional encoders. A counterfactual ablation makes
the scope honest: a full-rank supervised probe on the same features
reaches AUC $0.911$, and projecting our axis \emph{out} changes that by
only $0.001$; valence is richly distributed, and our axis is the most
useful one-dimensional summary extractable under a zero-label
constraint. The method is zero-label, nine-language, one-axis,
seventeen-evaluation, and \speedup{} faster than direct prompting.
\end{abstract}
