\begin{abstract}
We introduce \ours, a remarkably simple modification to standard deep
ensembling: train the same architecture with the same random seed and the
same cosine learning-rate schedule for two different lengths (e.g., 100 vs.\
150 epochs), and ensemble the two resulting checkpoints. The two training
trajectories reach \emph{identical} mean validation accuracy (differing by
$<\!0.001$ in our experiments), but they differ on 13\% of test samples
because the longer trajectory spends 50 additional epochs in a low-LR
``basin-exploration'' phase that drifts weights into a neighboring local
minimum. Averaging softmax probabilities across both trajectories produces a
+1.1 to +1.5 percentage-point gain \emph{beyond} what standard same-length
deep ensembling (5 or 10 seeds) achieves.

We validate \ours across three modalities (EEG, images, text) and nine
benchmarks spanning four architecture families. In EEG emotion
classification, we achieve a new state of the art on FACED 9-class
(\facedacc{} balanced accuracy, +5.3 pp / +8.2\% over the EMOD AAAI 2026
baseline, and +1.1 pp above a true 10-seed same-length deep ensemble),
matched by a cross-dataset confirmation on SEED-V (BAcc 0.4083,
\seedvgain{} ensemble gain). In image classification, we replicate the
trick on CIFAR-10, CIFAR-100, MNIST, and Fashion-MNIST with small CNN,
ResNet-18, and MLP backbones. In text classification, we confirm the
gain with DistilBERT on SST-2. Head-to-head against SWA, snapshot
ensembles, and standard 10-seed deep ensembles, \ours{} wins on all
four calibration metrics (accuracy, NLL, ECE, Brier) while adding
\emph{zero} hyperparameter tuning and a single additional training run.

A theoretical decomposition shows that \ours-induced diversity is
proportional to the ``unused'' individual-accuracy headroom: across nine
benchmarks the gain obeys a power-law in headroom
($\alpha\!=\!0.85$, Pearson $r\!=\!0.75$, $p\!=\!0.019$, $R^2\!=\!0.57$
for a linear fit), ranging from \facedgain{} on FACED (individual mean
0.66) down to +0.14\% on MNIST (individual mean 0.985). The trick is most
valuable where ensembling is most needed---on capacity-bound students
near their effective ceiling.
\end{abstract}
